### 2019-05-28 initial/notes.tex
check in some new equations + plots to notes

\subsection{Cassini State Hamilttonian}

@@

>>>>

\subsection{Cassini State Hamiltonian}

@@

\section{Weak Tidal Friction, Take 2}

I did a bunch of work w/o writing it up, so this is the basic gist.

\begin{itemize}
    \item We can still reuse the same equations as earlier, and the key point to
        notice is that for a given trajectory, only times $\phi = 0$ are
        important for determining the final fate of the system. Call these
        $\mu_0$, then the dynamics of the system can be described by a map
        $\mu_{0, i + 1}\p*{\mu_{0, i}, s}$. The reason $\mu_0$ is important is
        that only when $\mu_0 = \mu_4(s)$ CS4 can separatrix hopping occur.
        Thus, whenever $\mu_0$ crosses $\mu_4(s)$, there is some probability of
        entering the separatrix and some probability of hopping over it.

    \item The dynamics of the system are then governed by the dynamics of the
        map near $\mu_4$. In the limit where $\phi$ is approximately constant
        over a precession ($\mu_0$ is far from $\mu_4$ compared to $\sqrt{\eta
        \sin I}$ the separatrix width), then the map obeys nearly the same
        dynamics as the continuous flow in $(\mu, s)$ space for tidal friction.

    \item However, when we are sufficiently close to $\mu_4$, we enter the
        $\mu_0 - \mu_4 \lesssim \sqrt{2\eta \sin I}$ regime. Here, let's note
        that the map takes on dynamics:
        \begin{align}
            \rd{\mu_0}{t} &= \pd{\mu_0}{H}\rd{H}{t},\nonumber\\
                &= \frac{\dot{\phi}}{\dot{\phi}(\phi = 0)}
                    \p*{\epsilon\p*{1 - \mu^2}\p*{\frac{2\Omega_1}{s}
                        - \mu}},\nonumber\\
            \ev*{\rd{\mu_0}{t}} &=
                \frac{1}{T}\int\limits_0^T
                    \frac{\dot{\phi}}{\dot{\phi}_0}
                        \epsilon\p*{1 - \mu^2}\p*{\frac{2\Omega_1}{s} - \mu}
                            \;\mathrm{d}t,\nonumber\\
                &= \frac{1}{T}\int\limits_0^{2\pi}
                    \frac{1}{\dot{\phi}_0}
                        \epsilon\p*{1 - \mu^2}\p*{\frac{2\Omega_1}{s} - \mu}
                            \;\mathrm{d}\phi,\nonumber\\
                &\sim \frac{\sqrt{\eta \cos I}}{2\pi}
                    \frac{1}{\dot{\phi}(\phi = 0)}
                    \int\limits_0^{2\pi}
                        \epsilon\p*{1 - \mu^2}\p*{\frac{2\Omega_1}{s} - \mu}
                        \;\mathrm{d}\phi.\label{eq:final_approx}
        \end{align}
        We have substituted $\cos I/2\pi \sim 1/T$ a rough estimate
        for the period. But now, $\frac{1}{\dot{\phi}(\phi = 0)}$ near $\mu_4$
        obeys the linearization about the saddle point, and importantly it is
        small. Near $\mu_4$, the linearization we computed above tells us that
        $\dot{\phi} \approx - \sin I \Delta \mu$, and so we can imagine the
        true map dynamics by a flow with $\rd{\mu_0}{t}$ of form
        \begin{equation}
            \rd{\mu_0}{t} =
                \begin{cases}
                    \rd{\mu_0}{t} & \abs*{\Delta \mu} \gtrsim
                        2\sqrt{\eta \sin I},\\
                    -\frac{\cot I}{2\pi \Delta \mu}
                        \epsilon\p*{1 - \mu_{eff}^2}
                            \p*{\frac{2\Omega_1}{s} - \mu_{eff}}
                            & \text{otherwise}.
                \end{cases}
        \end{equation}
        We've defined $\Delta \mu = \mu_0 - \mu_4$ for convenience. $\mu_{eff}$
        reflects the fact that the $\epsilon$ integral in
        \autoref{eq:final_approx} is taken over a curve of very non-constant
        $\mu$, roughly following the separatrix boundary. In fact, $\mu_{eff}
        \sim \sqrt{2\eta \sin I}$ the separatrix width, likely.

        Our approximation near $\mu_4$ is probably only accurate up to scaling,
        we should ensure this piecewise definition is smooth. Thus, we choose
        $\mu_{eff} \equiv \mu_0 + 2\sqrt{\eta \sin I}$ and change the close-in
        term to be $\frac{\Delta \mu_{eff}}{\Delta
        \mu}\at{\rd{\mu_0}{t}}_{\mu_{eff}}$.

        We can compare these predictions qualitatively to the actual integrated
        map. By integrating the true $(\mu, \phi, s)$ system over a period, we
        can derive an effective $\frac{\Delta \mu}{\Delta t}$ and compare to the
        scaling we obtain above. The coefficients are not perfect, but the
        $1/\Delta \mu$ scaling is correct: \autoref{fig:pw}.
        \begin{figure}[t]
            \centering
            \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{1_weaktide/3diags0_7.png}
                \caption{$\frac{\Delta \mu}{\Delta t}$ computed using the pure
                weak tide prescription (ignoring $\mu(\phi)$ variations),
                computed using an integral of the full system and computed using
                the piecewise definition above.}
            \end{subfigure}
            \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{1_weaktide/3diags_loglog0_7.png}
                \caption{Loglog plot of $\Delta \mu/\Delta t$ above $\mu_4$ for
                the integrated and piecewise solutions.}
            \end{subfigure}
            \caption{Agreement of piecewise with integrated,
            qualitatively.}\label{fig:pw}
        \end{figure}

    \item The thing that is interesting now is that $\mu_{eff}$ is evaluated
        relatively farther away from $\mu_4$, and since $\rd{\mu}{t}$ in the
        weak tidal limit for $s > 2$ can be negative, it is possible for a point
        above $\mu_4$ to be \emph{sucked in} to CS4 \emph{from both above and
        below}. This is an extremely important feature, and indeed in
        \autoref{fig:pw} we see such behavior. If we are in this regime, then
        any nonzero separatrix hopping probability will drive 100\% of points
        incident on CS4 to enter the separatrix, since repeated hoppings must be
        observed!

        Qualitatively, we can use our piecewise definition to see the behavior
        of this transition; it occurs when $\rd{\mu}{t}$ evaulated at the
        $\mu_{eff}$ above $\mu_4$ is negative, i.e.\ above
        $\frac{2\Omega_1}{s}$. Thus, this boils down to $\mu_4 + 2\sqrt{\eta
        \sin I} \gtrsim \frac{2\Omega_1}{s}$. Since $\mu_4 \approx \eta \cos I
        = \frac{s_c\cos I}{s}$, we see that for smaller $s_c$ than some critical
        threshold, $\mu_4$ is no longer attracting on both sides. We can
        numerically solve for this transition using the full integrated
        equations: \autoref{fig:3crits}.
        \begin{figure}[t]
            \centering
            \includegraphics[width=0.6\textwidth]{1_weaktide/3crits.png}
            \caption{Critical values of $s_c$ below which $\mu_4$ is no longer
            strongly attracting.}\label{fig:3crits}
        \end{figure}

    \item Indeed, \autoref{fig:3crits} shows that the critical $s_c$ is a fairly
        flat function of $s$ in our parameter regime of interest. Call this
        critical $s_c$ $S_0$ owing to unfortunate notation. We can thus identify
        two regimes:
        \begin{itemize}
            \item $s_c < S_{0}$: points above CS4 will be driven away from CS4,
                and so only points below the separatrix will experience a single
                chance to separatrix hop. Thus, in this regime, points initially
                above the separatrix will tidally synchronize, while points
                below the separatrix will probabilistically either synchronize
                or go to CS2.

            \item $s_c > S_0$: CS4 becomes \emph{attracting}, and so two
                consequences: points initially above the separatrix are sucked
                downwards, and all encounters go to CS2 after 100\% separatrix
                entry probability.
        \end{itemize}
        Since $S_0$ is some function of $s$, it is clear that for intermediate
        values of $s_C$, we will see single separatrix hopping at early times
        while separatrix attraction at late times.
\end{itemize}
